---
title: "Assignment 3"
editor_options: 
  chunk_output_type: console
---

## Setup 

Load the required packages and read in the dataset below. 

```{r}
#| output: false
library(tidyverse)
library(boot)
library(ggthemes)

data_ai <- readr::read_csv("https://alopilato88.github.io/quantitative-analysis-1/assignments/03-assignment-data.csv")
```

## Question 1

**Q1: Create a table that shows what proportion of employees belong to a given a region. Which region do most employees work?**

```{r}

```

## Question 2

**Q2: Estimate a regression model that uses `office_region` to predict `perceived_freq_use`. Name this model mod_freq_use_1**

```{r}

```

## Question 3

**Q3: Does an individual's `perceived_freq_use` differ by `office_region`? Provide a brief explanation.**

```{r}

```

## Question 4

**Q4: Using `mutate`, transform the `office_region` variable into a factor and make the reference group equal to `north america`. Name this new variable: `office_region_factor`.**

```{r}

```

## Question 5

**Q5: Estimate a regression model that uses `office_region_factor` to predict `perceived_freq_use`. Name this model: mod_freq_use_2.**

```{r}

```

## Question 6

**Q6: Why do the regression coefficients in `mod_freq_use_1` differ from the coefficients in `mod_freq_use_2`?**

```{r}

```

## Question 7

**Q7: Estimate a regression model that tests if `previous_exp` moderates the relationship between `perceived_useful` and `behavioral_intention`. In this model, `perceived_useful` is the focal predictor and `previous_exp` is the moderator. Name the model `mod_beh_intent_1`.** 

```{r}

```

## Question 8

**Q8: Does `previous_exp` moderate the relationship between `perceived_useful` and `behavioral_intention`? Why or why not?**

```{r}

```

## Question 9

**Q9: Interpret the interaction you estimated in `mod_beh_intent_1`. Feel free to use a plot using `ggplot2` to aid in your interpretation.**

```{r}

```

## Question 10

**Q10: Is the conditional relationship between `perceived_useful` and `behavioral_intention` for employees in the `prev_exp == yes` group? What about for the `prev_exp == no` group? Explain.** 

```{r}

```

## Question 11

**Q11: Estimate several regression models that you can use to determine if `perceived_useful` indirectly affects `perceived_freq_use` through `behavioral_intention`. Name the model with `behavioral_intention` as the outcome: `mod_ie` and name the model with `perceived_freq_use` as the outcome: `mod_de`.**

```{r}

```

## Question 12

**Q12: In the models you fit in Question 11, which variable is the antecedent variable, which variable is the mediator, and which variable is the outcome?**


## Question 13

**Q13: What the direct effect of `perceived_useful` on `behavioral_intention`? What is the indirect effect of `perceived_useful` on `behavioral_intention`? Calculate the total effect using the indirect effect and direct effect, then compare it to the estimated total effect.**

```{r}

```

## Bonus Question

**Use the bootstrap method to determine if the indirect effect is significant. Use 500 bootstrap samples. Use `set.seed(8547)`, so  that I can reproduce your results.**

```{r}

```


